---
title: 'Used Cars for Sale'
output:
  pdf_document: default
---

# Introduction  
|       This project explores a dataset about cars for sale taken from the website [TrueCar.com](https://www.truecar.com/) on September 24^th^, 2017. The dataset, which I downloaded from [kaggle.com](https://www.kaggle.com/jpayne/852k-used-car-listings?select=tc20171021.csv), includes roughly 1.2 million rows of cars' make, model, year, mileage, listing price, VIN, and the city and state in which the car was for sale. 

|       While the original purpose of this investigation was to determine which cars were worth the price tag, it quickly turned into an exploration of determining model popularity, trends of the seller's location, average age, average mileage, average price of models, and the distribution of electric vehicles throughout the country and state. By studying this data, consumers can inform themselves on which type of vehicle might be best fit for the area in which they live, how rare a model they're looking for might be, finding which variation of models are the least expensive, give them a general idea about the average price, age, and mileage of cars they're interested in, and more. Dealerships may also be able to benefit from this data by finding which models and variations are most popular as well as what areas need more cars than others. 


# Load Packages 
|       Each of the following packages allows me to easily work with the data and create graphics that display my findings in an interesting, informative, and effective way. I added the "options(scipen = 999)" to ensure that my data did not to revert to scientific notation.  

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(error = TRUE)
options(tigris_use_cache = TRUE)

library(tidyverse)
library(plotly)
library(treemapify)
library(tigris)
library(forcats)
library(maps)
library(cowplot)
options(scipen = 999)

```


# Load the Data 
|     The following datasets were loaded into RStudio to complete my project. The cars dataset is the one aforementioned, while the regions dataset includes the full names of the states in the US, their abbreviations, and which region the state is a part of (south, west, northeast, or midwest).  

```{r}
cars <- read_csv("tc20171021.csv")
regions <- read_rds("states.rds")
```


# Tidy the Data
|       Since the dataset I wanted to use was so large, I decided to limit my project to focus only on the top ten most popular manufacturers. In descending order, this turns out to be: Ford, Chevy, Toyota, Nissan, Honda, Jeep, Hyundai, Dodge, GMC, and Kia. After narrowing down my data, one of the biggest hurdles I faced was determining how to clean it up. I had many rows where the model type was grouped with a descriptor. For example, there were cars listed as a Nissan Altima4dr, or a Ford F-150FWD. While this information did qualify as the car's model, I wanted the model information to be split into two separate columns: model and and model variation. This way, I could search through Nissan Altimas or Ford F-150s and not worry about how many doors it had, what type of drive it was, whether it was luxury or not, etc. 

|       To fix this issue, I used both RStudio and Excel. To create a tidy dataset for each make I had to manually input how many letters were in each model's name. In these new datasets I also changed the missing model variations become _NA_.  

```{r}
cars1 <- cars %>%
  filter(Make == "Ford")

cars2 <- cars1 %>%
  mutate(Model_part = str_sub(Model, 1, 7)) %>%
  group_by(Make, Model_part) %>%
  summarise(Models = str_c(unique(Model), collapse = ","))

write_csv(cars2, "cars2Ford.csv")

cars3 <- read_csv("Tidying the Data Files/cars2Ford_first_word (1).csv")

cars4 <- cars3 %>%
  separate_rows(Models, sep = ",") %>%
  select(Model= Models, first_word)

cars5 <- cars1 %>%
  left_join(cars4, by = "Model")

Ford <- cars5 %>%
  mutate(
    Model_Type = str_sub(Model, 1, first_word),
    Model_Variation = str_sub(Model, (first_word+1), -1)
  )
Ford[Ford == ""] <- NA
```

```{r}
cars1 <- cars %>%
  filter(Make == "Chevrolet")

cars2 <- cars1 %>%
  mutate(Model_part = str_sub(Model, 1, 4)) %>%
  group_by(Make, Model_part) %>%
  summarise(Models = str_c(unique(Model), collapse = ","))

write_csv(cars2, "cars2Chevy.csv")

cars3 <- read_csv("Tidying the Data Files/cars2Chevy_first_word.csv")

cars4 <- cars3 %>%
  separate_rows(Models, sep = ",") %>%
  select(Model= Models, first_word)

cars5 <- cars1 %>%
  left_join(cars4, by = "Model")

Chevy <- cars5 %>%
  mutate(
    Model_Type = str_sub(Model, 1, first_word),
    Model_Variation = str_sub(Model, (first_word+1), -1)
  )
Chevy[Chevy == ""] <- NA
```

```{r}
cars1 <- cars %>%
  filter(Make == "Toyota")

cars2 <- cars1 %>%
  mutate(Model_part = str_sub(Model, 1, 4)) %>%
  group_by(Make, Model_part) %>%
  summarise(Models = str_c(unique(Model), collapse = ","))

write_csv(cars2, "cars2Toyota.csv")

cars3 <- read_csv("Tidying the Data Files/cars2Toyota_first_word.csv")

cars4 <- cars3 %>%
  separate_rows(Models, sep = ",") %>%
  select(Model= Models, first_word)

cars5 <- cars1 %>%
  left_join(cars4, by = "Model")

Toyota <- cars5 %>%
  mutate(
    Model_Type = str_sub(Model, 1, first_word),
    Model_Variation = str_sub(Model, (first_word+1), -1)
  )
Toyota[Toyota == ""] <- NA
```

```{r}
cars1 <- cars %>%
  filter(Make == "Nissan")

cars2 <- cars1 %>%
  mutate(Model_part = str_sub(Model, 1, 3)) %>%
  group_by(Make, Model_part) %>%
  summarise(Models = str_c(unique(Model), collapse = ","))

write_csv(cars2, "cars2Nissan.csv")

cars3 <- read_csv("Tidying the Data Files/cars2TNissan_first_word.csv")

cars4 <- cars3 %>%
  separate_rows(Models, sep = ",") %>%
  select(Model= Models, first_word)

cars5 <- cars1 %>%
  left_join(cars4, by = "Model")

Nissan <- cars5 %>%
  mutate(
    Model_Type = str_sub(Model, 1, first_word),
    Model_Variation = str_sub(Model, (first_word+1), -1)
  )
Nissan[Nissan == ""] <- NA
```

```{r}
cars1 <- cars %>%
  filter(Make == "Honda")

cars2 <- cars1 %>%
  mutate(Model_part = str_sub(Model, 1, 3)) %>%
  group_by(Make, Model_part) %>%
  summarise(Models = str_c(unique(Model), collapse = ","))

write_csv(cars2, "cars2Honda.csv")

cars3 <- read_csv("Tidying the Data Files/cars2Honda_first_word.csv")

cars4 <- cars3 %>%
  separate_rows(Models, sep = ",") %>%
  select(Model= Models, first_word)

cars5 <- cars1 %>%
  left_join(cars4, by = "Model")

Honda <- cars5 %>%
  mutate(
    Model_Type = str_sub(Model, 1, first_word),
    Model_Variation = str_sub(Model, (first_word+1), -1)
  )
Honda[Honda == ""] <- NA
```

```{r}
cars1 <- cars %>%
  filter(Make == "Jeep")

cars2 <- cars1 %>%
  mutate(Model_part = str_sub(Model, 1, 4)) %>%
  group_by(Make, Model_part) %>%
  summarise(Models = str_c(unique(Model), collapse = ","))

write_csv(cars2, "cars2Jeep.csv")

cars3 <- read_csv("Tidying the Data Files/cars2Jeep_first_word.csv")

cars4 <- cars3 %>%
  separate_rows(Models, sep = ",") %>%
  select(Model= Models, first_word)

cars5 <- cars1 %>%
  left_join(cars4, by = "Model")

Jeep <- cars5 %>%
  mutate(
    Model_Type = str_sub(Model, 1, first_word),
    Model_Variation = str_sub(Model, (first_word+1), -1)
  )
Jeep[Jeep == ""] <- NA
```

```{r}
cars1 <- cars %>%
  filter(Make == "Hyundai")

cars2 <- cars1 %>%
  mutate(Model_part = str_sub(Model, 1, 3)) %>%
  group_by(Make, Model_part) %>%
  summarise(Models = str_c(unique(Model), collapse = ","))

write_csv(cars2, "cars2Hyundai.csv")

cars3 <- read_csv("Tidying the Data Files/cars2Hyundai_first_word.csv")

cars4 <- cars3 %>%
  separate_rows(Models, sep = ",") %>%
  select(Model= Models, first_word)

cars5 <- cars1 %>%
  left_join(cars4, by = "Model")

Hyundai <- cars5 %>%
  mutate(
    Model_Type = str_sub(Model, 1, first_word),
    Model_Variation = str_sub(Model, (first_word+1), -1)
  )
Hyundai[Hyundai == ""] <- NA
```

```{r}
cars1 <- cars %>%
  filter(Make == "Dodge")

cars2 <- cars1 %>%
  mutate(Model_part = str_sub(Model, 1, 4)) %>%
  group_by(Make, Model_part) %>%
  summarise(Models = str_c(unique(Model), collapse = ","))

write_csv(cars2, "cars2Dodge.csv")

cars3 <- read_csv("Tidying the Data Files/cars2Dodge_first_word.csv")

cars4 <- cars3 %>%
  separate_rows(Models, sep = ",") %>%
  select(Model= Models, first_word)

cars5 <- cars1 %>%
  left_join(cars4, by = "Model")

Dodge <- cars5 %>%
  mutate(
    Model_Type = str_sub(Model, 1, first_word),
    Model_Variation = str_sub(Model, (first_word+1), -1)
  )
Dodge[Dodge == ""] <- NA
```

```{r}
cars1 <- cars %>%
  filter(Make == "GMC")

cars2 <- cars1 %>%
  mutate(Model_part = str_sub(Model, 1, 3)) %>%
  group_by(Make, Model_part) %>%
  summarise(Models = str_c(unique(Model), collapse = ","))

write_csv(cars2, "cars2GMC.csv")

cars3 <- read_csv("Tidying the Data Files/cars2GMC_first_word.csv")

cars4 <- cars3 %>%
  separate_rows(Models, sep = ",") %>%
  select(Model= Models, first_word)

cars5 <- cars1 %>%
  left_join(cars4, by = "Model")

GMC <- cars5 %>%
  mutate(
    Model_Type = str_sub(Model, 1, first_word),
    Model_Variation = str_sub(Model, (first_word+1), -1)
  )
GMC[GMC == ""] <- NA
```

```{r}
cars1 <- cars %>%
  filter(Make == "Kia")

cars2 <- cars1 %>%
  mutate(Model_part = str_sub(Model, 1, 3)) %>%
  group_by(Make, Model_part) %>%
  summarise(Models = str_c(unique(Model), collapse = ","))

write_csv(cars2, "cars2Kia.csv")

cars3 <- read_csv("Tidying the Data Files/cars2Kia_first_word.csv")

cars4 <- cars3 %>%
  separate_rows(Models, sep = ",") %>%
  select(Model= Models, first_word)

cars5 <- cars1 %>%
  left_join(cars4, by = "Model")

Kia <- cars5 %>%
  mutate(
    Model_Type = str_sub(Model, 1, first_word),
    Model_Variation = str_sub(Model, (first_word+1), -1)
  )
Kia[Kia == ""] <- NA
```

|       Once this process was completed for each of the top ten most popular makes, I wanted to combine all of the tidied datasets to create a comprehensive one called "pretty_cars". I also wanted this pretty_cars dataset to have the region in which the car's posting was, so I joined pretty_cars to the regions dataset. There was also an issue with the inconsistency of the capitalization in the state abbreviations, so I changed the new dataset to have all uppercase state abbreviations.  

```{r}
pretty_cars <- rbind.data.frame(Ford, Chevy, Toyota, Nissan, Honda,
                                Jeep, Hyundai, Dodge, GMC, Kia) 

pretty_cars <- pretty_cars %>%
  left_join(regions, by = c("State" = "state")) 

pretty_cars <- pretty_cars %>%
  mutate(State = str_to_upper(State)) 

head(pretty_cars)
```

|       The pretty_cars dataset now has the original "Model" column as well as the new "Model_Type" and "Model_Variation". You can also see the "first_word" column represents how many letters are in the model. The regional information has also been added.

# Analyze the Data
## Popularity

|       As previously mentioned, I determined the top ten most popular makes so that I could narrow down the data I would be working with for the rest of the report. Once I established the "pretty_cars" dataset, I wanted to get a glimpse into each of the manufacturer's popularity over the years and approximately how many cars where produced each year. 

```{r}
pretty_cars %>%
  ggplot(aes(Year, fill = Make))+
  geom_bar()+
  theme_bw()+
  scale_y_continuous(expand = c(0,0))+
  labs(title = "Car Popularity Each Year", y = "Number of Cars")+
  theme(plot.title = element_text(hjust = .5))
```

|       From this, we begin to see that the ages of the cars for sale ranged from 20 years old to brand new (since the dataset was taken in 2017). Expanding upon that idea, I was interested to find which models were most popular within each make. To show this, I created a tree map where the area of the boxes represents the amount of cars in a make/model.  

```{r}
pretty_cars %>%
  count(Make, Model_Type) %>%
  ggplot(aes(area = n, fill = Make, subgroup = Make))+
  geom_treemap(aes(alpha = n/100))+
  geom_treemap_text(color = "white", alpha = .5, aes(label = Model_Type))+
  geom_treemap_subgroup_text(place = "center", color = "white")+
  geom_treemap_subgroup_border(color = "white")+
  theme(legend.position = "none")
```

|       From these graphs, it was interesting to see that Ford, Chevy, and Toyota accounted for nearly half of the dataset themselves. Within the Ford and Chevy makes, F-150s and Silverados were by far the most popular models. The graphs also showed that Mustangs and Camaros were decently popular models for Ford and Chevy, which is not something I expected to see.

|       I also wanted to visualize what variations were most popular within each model for Fords, Chevys, and Toyotas. I found that there was an excessive amount of model variation options, so I limited the graphs to show only the top 4 most popular variations with the remaining variations grouped in an "Other" category. Since there were a lot of different model types within each make, I also filtered the graph to only show the models which had more than 100 appearances in the data.  

```{r}
pretty_cars %>%
  filter(Make == "Ford") %>%
  group_by(Make) %>%
  mutate(adjusted_model_variation = fct_lump_n(Model_Variation, 4)) %>%
  count(Model, Model_Type, Model_Variation, adjusted_model_variation) %>%
  filter(n > 100) %>%
  ggplot(aes( x = reorder(Model_Type, -n, sum), y = n,  fill = adjusted_model_variation))+
  geom_col()+
  theme_bw()+
  scale_y_continuous(expand = c(0,0))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .3), 
        plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(hjust = .5),
        plot.caption = element_text(size = 8))+
  labs(title = "Ford Model and Variation Popularity", x = "Model", y = "Number of Cars", 
       fill = "Model Variation", caption = "For Model Counts Greater than 100")
```

```{r}
pretty_cars %>%
  filter(Make == "Chevrolet")%>%
  group_by(Make) %>%
  mutate(adjusted_model_variation = fct_lump_n(Model_Variation, 4)) %>%
  count(Model, Model_Type, Model_Variation, adjusted_model_variation) %>%
  filter(n > 100) %>%
  ggplot(aes( x = reorder(Model_Type, -n, sum), y = n,  fill = adjusted_model_variation))+
  geom_col()+
  theme_bw()+
  scale_y_continuous(expand = c(0,0))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .3), 
        plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(hjust = .5),
        plot.caption = element_text(size = 8))+
  labs(title = "Chevy Model and Variation Popularity", x = "Model", y = "Number of Cars",
       fill = "Model Variation", caption = "For Model Counts Greater than 100")
```

```{r}
pretty_cars %>%
  filter(Make == "Toyota") %>%
  group_by(Make) %>%
  mutate(adjusted_model_variation = fct_lump_n(Model_Variation, 4)) %>%
  count(Model, Model_Type, Model_Variation, adjusted_model_variation) %>%
  filter(n > 100) %>%
  ggplot(aes( x = reorder(Model_Type, -n, sum), y = n,  fill = adjusted_model_variation))+
  geom_col()+
  theme_bw()+
  scale_y_continuous(expand = c(0,0))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .3), 
        plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(hjust = .5),
        plot.caption = element_text(size = 8))+
  labs(title = "Toyota Model and Variation Popularity", x = "Model", y = "Number of Cars",
       fill = "Model Variation", caption = "For Model Counts Greater than 100")
```

|       From this we can see that Ford F-150s had many different kinds of model variations while Chevy Silverados had none. I did continue to make these graphs for the remaining 7 makes and found that Honda Accords, Jeep Wranglers, Jeep Grand Cherokees, Hyundai Santa Fes, Dodge Grand Caravan, Dodge Rams, and GMC Sierras had nearly no variations within the models. There was also a pattern that the type of drive (4WD, 2WD, AWD, FWD), the number of doors (4dr, 2dr), and the luxury level of the vehicles (SE, LT, LX, sport, etc.) accounted for most of the model variations. As a whole, it seemed as though larger vehicles (trucks and SUVs) are more popular than smaller vehicles. 

## Location 

|       Location wasn't a factor that I was originally concerned with, but it became more prevalent as I began finding trends within the data. Because of this, I wanted to see the distribution of models across the US and within each state and region. Since Fords, Chevys, and Toyotas were the most popular makes, I created to following graphs to show their presence in the United States.  

```{r}
ford_state_data <- pretty_cars %>%
  filter(Make == "Ford") %>%
  filter(!(State == "DC")) %>%
  group_by(State) %>%
  count(State)

us_states <- states(year = 2019, resolution = "20m")

ford_us_state_cont <- us_states %>%
  filter(!STATEFP %in% c("15", "78", "69", "66", "02", "60", "72")) %>%
  left_join(ford_state_data, by = c("STUSPS" = "State")) 

ggplot()+
  geom_sf(
    data = ford_us_state_cont,
    aes(fill = n),
    color = "black",
    size = .2)+
  theme_map()+
  labs(title = "Popularity of Fords in the United States", fill = "Number of Cars")+
  theme(plot.title = element_text(hjust = .5))+
  scale_fill_gradient(low = "#56B1F7", high = "#132B43")
```

```{r}
chevy_state_data <- pretty_cars %>%
  filter(Make == "Chevrolet") %>%
  filter(!(State == "DC")) %>%
  group_by(State) %>%
  count(State)

us_states <- states(year = 2019, resolution = "20m")

chevy_us_state_cont <- us_states %>%
  filter(!STATEFP %in% c("15", "78", "69", "66", "02", "60", "72")) %>%
  left_join(chevy_state_data, by = c("STUSPS" = "State")) 

ggplot()+
  geom_sf(
    data = chevy_us_state_cont,
    aes(fill = n),
    color = "black",
    size = .2)+
  theme_map()+
  labs(title = "Popularity of Chevys in the United States", fill = "Number of Cars")+
  theme(plot.title = element_text(hjust = .5))+
  scale_fill_gradient(low = "#56B1F7", high = "#132B43")
```

```{r}
toyota_state_data <- pretty_cars %>%
  filter(Make == "Toyota") %>%
  filter(!(State == "DC")) %>%
  group_by(State) %>%
  count(State)

us_states <- states(year = 2019, resolution = "20m")

toyota_us_state_cont <- us_states %>%
  filter(!STATEFP %in% c("15", "78", "69", "66", "02", "60", "72")) %>%
  left_join(toyota_state_data, by = c("STUSPS" = "State")) 

ggplot()+
  geom_sf(
    data = toyota_us_state_cont,
    aes(fill = n),
    color = "black",
    size = .2)+
  theme_map()+
  labs(title = "Popularity of Toyotas in the United States", fill = "Number of Cars")+
  theme(plot.title = element_text(hjust = .5))+
  scale_fill_gradient(low = "#56B1F7", high = "#132B43")
```

|       I also created a visual to see which models were deemed the most popular in each state. I also wanted to determine regional popularity, so I made a chart showing the top 5 models for each region.

```{r}
state_data <- pretty_cars %>%
  filter(!(State == "DC")) %>%
  group_by(State) %>%
  count(Model_Type, Make, sort = TRUE) %>%
  slice_max(n)

us_states <- states(year = 2019, resolution = "20m")

us_state_cont <- us_states %>%
  filter(!STATEFP %in% c("15", "78", "69", "66", "02", "60", "72")) %>%
  left_join(state_data, by = c("STUSPS" = "State")) %>%
  filter(!(is.na(Model_Type)))

ggplot()+
  geom_sf(
    data = us_state_cont,
    aes(fill = Model_Type),
    color = "grey48",
    size = .2)+
  theme_map()+
  scale_fill_manual(values = c( "midnightblue", "mediumpurple4", "darkorchid4","orchid3",
                                "royalblue3", "lightslateblue","blueviolet", "violet",
                                "plum2"),name = "Model")+
  labs(title = "Most Popular Models in the United States")+
  theme(plot.title = element_text(hjust = .5))
```

```{r}
region_popularity_colors <- c("Accord" = "blue4", "Civic" = "mediumpurple3", 
                              "CR-V" = "pink2", "Altima" = "mediumorchid",
                              "Versa" = "grey77","Camry" = "cornflowerblue",
                              "Escape" = "deeppink3", "F-150" = "magenta4",
                              "Grand" = "plum", "Malibu" = "darkslateblue",
                              "Silverado" = "slateblue2")

pretty_cars %>%
  group_by(region) %>%
  count(Model_Type, sort = TRUE) %>%
  slice(1:5) %>%
  ggplot(aes( x = reorder(region, -n), y = n, fill = Model_Type))+
  geom_bar(position = "dodge", stat = "identity")+
  scale_y_continuous(expand = c(0,0))+
  labs(title = "Model Popularity within Regions", x = "Number of Cars", y = "Region",
       fill = "Model", caption = "Where NA accounts for Alaska and DC")+
  scale_fill_manual(values = region_popularity_colors)+
  theme_bw()+
  theme(plot.title = element_text(hjust = .5), plot.caption = element_text(size = 8))
```

|       Evidently, the distribution of Fords, Chevys, and Toyotas show that there is an abundance of cars in Texas, California, and Florida. Each of the makes seem to be equivalent in their distribution across the country, and the regional graph reassures that there are more cars in the south than any other region. Since the regional bar graph accounts for the 5 most popular models in each region, we can see that F-150s and Silverados are the most popular types of vehicles in every region except the northeast, interestingly enough. In the states from Pennsylvania up to Maine, Honda Accords and Jeep Grand Cherokees are the most popular vehicles. In fact, Ford F-150s aren't in the top 5 models for the northeast region at all. 

|       The map of the most popular model for each state shows again that F-150s and Silverados are the most popular model in a majority of the states. In the northeast region, you can also verify that Accords are the most popular. California, Florida, Virginia, Delaware, Massachusetts, and Illinois seem to be outliers in the fact that their most popular vehicle is unique from the rest of the country. In these states, the most popular vehicle is a type of car while the rest of the states have SUVs or trucks as their most popular vehicles. 


## Mileage, Age, and Price 
|       One of the main details that I wanted to explore within the dataset were the trends among age, mileage, and price. In addition to finding the averages of these factors for each model, I also wanted to compare this information with the averages of the most popular variation within each model.  

```{r}
most_popular_model_averages <- pretty_cars %>%
  filter(Model_Type == c("F-150", "Silverado", "Camry", "Altima", "Accord", "Sonata",
                         "Sierra", "Optima", "Wrangler")|
           Make == "Dodge" & Model_Type == "Grand") %>%
  group_by(Model_Type) %>%
  mutate(average_year = mean(Year)) %>%
  mutate(average_mileage = mean(Mileage)) %>%
  mutate(average_price = mean(Price)) %>%
  slice(1)



pretty_cars %>%
  filter(Model_Type == "F-150" & Model_Variation == "4WD" | 
           Model_Type == "Silverado" & is.na(Model_Variation) | 
           Model_Type == "Camry" & Model_Variation == "SE"|
           Model_Type == "Altima" & Model_Variation == "2.5"|
           Model_Type == "Accord" & is.na(Model_Variation) |
           Model_Type == "Wrangler" & is.na(Model_Variation) |
           Model_Type == "Sonata" & Model_Variation == "2.4L" |
           Make == "Dodge" & Model_Type == "Grand" & is.na(Model_Variation) |
           Model_Type == "Sierra" & is.na(Model_Variation) |
           Model_Type == "Optima" & Model_Variation == "LX") %>%
  group_by(Model_Type) %>%
  mutate(average_mileage = mean(Mileage)) %>%
  mutate(average_price = mean(Price)) %>%
  slice(1) %>%
  select(Make, Model_Type, Model_Variation, average_mileage, average_price) %>%
  arrange(desc(average_mileage)) %>%
  ggplot(aes(average_mileage, average_price,color = Model_Type))+
  geom_point()+
  scale_color_discrete(labels = c("Honda Accord NA", "Nissan Altima 2.5", "Toyota Camry SE", 
                                  "Ford F-150 4WD", "Dodge Grand Caravan NA", "Kia Optima LX",
                                  "GMC Sierra NA", "Chevy Silverado NA", "Hyundai Sonata 2.4L",
                                  "Jeep Wrangler NA"))+
  theme_bw()+
  labs(title = "Average Mileage and Price", 
       subtitle = "10 Most Popular Models and their Most Popular Variations",
       y = "Average Price", x = "Average Mileage", 
       caption = "Where Points Represent Variation Averages and Triangles Represent Model Averages",
       color = "Make and Model")+
  scale_y_continuous(labels = scales::dollar_format())+
  theme(plot.title = element_text(hjust = .5, size = 14), 
        plot.subtitle = element_text(hjust = .5, size = 8), 
        plot.caption = element_text(size = 7))+
  geom_point(data = most_popular_model_averages, aes(average_mileage,average_price), shape = 2)
```

|       It is in this graph that we can see how the most popular variation of a model compares to the model as a whole. Specifically, Nissan Altima 2.5s tend to have less mileage on them than Nissan Altimas as a whole. The same is true for Kia Optima LXs, Hyundai Sonata 2.4Ls, Toyota Camry SEs, and Jeep Wrangler NAs. You can also notice that the average mileage and price of Honda Accord NAs, Dodge Grand Caravan NAs, and Chevy Silverado NAs are close to the averages of all Honda Accords, all Dodge Grand Caravans, and all Chevy Silverados, respectively. Another trend this graph shows is how Ford F-150s, Chevy Silverados, GMC Sierras, and Jeep Wranglers have high prices for their mileage compared to the rest of the included vehicles. 

|       To continue in the exploration of model types and their average mileage, price, and year, I had the following graph display the averages of the model variations for the most popular models of the top 5 makes.  

```{r}
pretty_cars %>%
  filter(Model_Type == "Silverado" |
           Model_Type == "Camry" |
           Model_Type == "F-150" |
           Model_Type == "Altima" |
           Model_Type == "Accord") %>%
  group_by(Make, Model_Type, Model_Variation) %>%
  summarise(average_variation_price = mean(Price), average_year = mean(Year), 
            average_mileage = mean(Mileage)) %>%
  arrange(desc(average_variation_price)) %>%
  ggplot(aes(average_mileage, average_variation_price, color = Model_Type))+
  geom_point()+
  geom_smooth(method = lm, linetype = "dashed", fullrange = TRUE, se = FALSE, size = .3)+
  theme_minimal()+
  theme(plot.caption = element_text(size = 8),
        plot.title = element_text(hjust = .5))+
  labs(x = "Average Mileage", y = "Average Price", title = "Average Mileage and Price by Make", 
       caption = "Where Points are Averages of Model Variations", color = "Model")
```

|       One important thing to notice in this graph is that Chevy Silverados don't have an average line because there is only one variation (and therefore not enough information to make a trendline). Here we can see that the average price of Ford F-150s per mile is much greater than the other models. There does seem to be an intersection point of all but one of the models when mileage is 150,000 miles and price is almost $0. This could indicate that people are uninterested in paying for or selling a car with 150,000 plus miles on it. Out of the models that are cars, the three models (Accord, Altima, Camry) have very similar trendline slopes and intercepts. It's also interesting to note that the trendlines of the cars seem to fit better than the trendlines of the trucks. 

|       Anyone who is devoted to finding a good deal on a car knows to pay attention to its mileage. In addition to its variation or age, mileage can indicate how much longer the car will run. To gauge how the age of Fords, Chevys, and Toyotas impacts their mileage, I created a graph that shows the trend of mileage over time.  

```{r}
pretty_cars %>%
  filter(Make == "Ford" | Make == "Chevrolet"|Make == "Toyota") %>%
  group_by(Make, Model_Type, Year, Mileage) %>%
  summarise(average_price = mean(Price), average_mileage = mean(Mileage), 
            average_year = mean(Year)) %>%
  slice(1) %>%
  ggplot(aes(average_year, average_mileage, fill = Make))+
  geom_col(position = position_dodge(width = .7))+
  scale_fill_manual(values = c("darkslateblue", "mediumpurple3", "plum"))+
  scale_y_continuous(expand = c(0,0))+
  theme_bw()+
  labs(title = "Average Year and Mileage", x = "Average Year", y = "Average Mileage")+
  theme(plot.title = element_text(hjust = .5))
```

|       Here we can see that Toyotas tend to have the highest mileage over time, followed by Fords then Chevys. While the mileage of each make tends to decline as the cars get younger, the trend of Toyota mileage seems to be nearly bimodal with a second peak (although not as large as the first) occuring near the year 2007. Perhaps 2007 Toyotas were built with better parts to make them last particularly longer than other manufacturing years. This being said, it seems that on average, Toyotas can run longer than Chevys or Fords.

## Electric Vehicles
|       I also wanted to investigate the presence of electric vehicles within the country and state. I had expected to see a greater amount of EVs in cities, but I was unsure about their popularity in suburban or rural settings. Because of this, I created a graph that would compare which state had the most EVs and another that would show which city in PA had the most EVs. 

|       The easiest way for me to compare the presence electric vehicles from location to location was to add a column to the pretty_cars dataset called "cars_per_state". This "cars_per_state" column counts how many cars were listed for sale in each state. With this column, I could calculate the percentage of EVs rather than just the raw number. I applied the same method to create the second graph but instead of finding the number of cars per state, I found the number of "cars_per_city".  

```{r}
cars_per_state <- pretty_cars %>%
  count(State) %>%
  mutate(cars_per_state = n) %>%
  select(State, cars_per_state)

pretty_cars <- pretty_cars %>%
  full_join(cars_per_state, by = "State")

pretty_cars %>%
  filter(Model_Variation == "Hybrid" |
           Model_Type == "Bolt" |
           Model_Type == "Niro" |
           Model_Type == "LEAF" |
           Model_Variation == "EV") %>%
  group_by(State) %>%
  count(Model_Type, Make, City, cars_per_state) %>%
  ungroup() %>%
  mutate(ev_percentage = 100*(n/cars_per_state)) %>%
  arrange(desc(ev_percentage)) %>%
  ggplot(aes(x = reorder(State, -ev_percentage, sum),
             y = ev_percentage, 
             fill = Model_Type))+
  geom_col()+
  theme_bw()+
  scale_y_continuous(expand = c(0,0))+
  labs(x = "State", y = "EV Percentage", fill = "Model",
       title = "State Electric Vehicle Presence")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .3),
        plot.title = element_text(hjust = .5))
```

```{r}
cars_per_city <- pretty_cars %>%
  count(City) %>%
  mutate(cars_per_city = n) %>%
  select(City, cars_per_city)

pretty_cars <- pretty_cars %>%
  full_join(cars_per_city, by = "City")

pretty_cars %>%
  filter(Model_Variation == "Hybrid" |
           Model_Type == "Bolt" |
           Model_Type == "Niro" |
           Model_Type == "LEAF" |
           Model_Variation == "EV",
         State == "PA") %>%
  group_by(City) %>%
  count(Model_Type, Make, City, cars_per_city) %>%
  mutate(ev_percentage = 100*(n/cars_per_city)) %>%
  arrange(desc(ev_percentage)) %>%
  ggplot(aes(x = fct_reorder(City, -ev_percentage, sum), y = ev_percentage, 
             fill = Model_Type))+
  geom_col()+
  theme_bw()+
  scale_y_continuous(expand = c(0,0))+
  labs(x = "State", y = "Electric Vehicle Percentage", fill = "Model", 
       title = "EV Percentage in PA Cities")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .3, size = 8),
        plot.title = element_text(hjust = .5))
```

|       The first thing that stands out from these graphs is that Nissan LEAFs seem to be the most popular EV in the United States while Ford Fusion Hybrids are the most popular EV in PA. Another notable part of these graphs is that the EV percentages in Oregon, Washington, and California are much more substantial than many of the other states. While EVs still only account for about 1% of the cars for sale in each of these three states, the impact of 1% in California compared to less than .1% in Alaska is impressive in relativity. I thought it was also interesting to see that PA is near the tail end of the graph even though we have two large cities (Philadelphia and Pittsburg). 

|       Just as it was interesting to see the relationship of EV presence between the states, it's also interesting to see the relationship between the cities in PA. While I expected Philadelphia and Pittsburgh to have the greatest percentage of EVs, it turns out that a town I had never heard of before, Coraopolis, had the highest EV percentage instead. Coraopolis seemed to be made entirely of Nissan LEAFs when most of the other cities in PA were composed of Ford Fusions or a combination of the EV models. Similar to the graph comparing the states, the PA graph shows a drastic difference between the highest and lowest percentage with a not-so-smooth gradient in between. I thought my hometown, Lancaster, would make an appearance on this graph since it is a decent sized city but apparently its EV count is not significant enough to show up.  


# Final Notes

|       Overall, the exploration of this dataset clearly revealed several different patterns that car owners can consider when shopping for a new car or selling an old one. There is a lot that this report covered and still more that it didn't (like seeing the distribution of cars in cities, repair costs, mpg, etc.), but I believe that the information contained in this write up can be useful to many people. From those that have little to no knowledge of cars to those that know much more than me, the exploration of this dataset can be useful in a multitude of ways. 

